\papertitle

### 2017-03-08 paper/paper.tex
Added CRLB to plots
Improved plot arrangement.
Added figures to paper.

@@

\section{Introduction}
\label{sec:intro}
%Additive synthesis using a sum of sinusoids plus noise is a powerful model for
%representing audio, allowing for the easy implementation of many manipulations
%such as time-stretching and transposition. In its most basic form, peaks in the
%short-time spectrum are identified as representing sinusoids and connected based
%on some criteria. As the phase evolution of the sinusoid is assumed linear over
%the analysis frame, only the phase and frequency of the sinusoids at these
%analysis points are used to fit a plausible phase function. Recently, there has
%been interest in using higher order phase functions as the estimation of their
%parameters has been made possible by a new set of techniques of only moderate
%computational complexity using signal derivatives. Theoretically, the estimation
%of the parameters of a polynomial phase function of arbitrary order make it
%possible to approximate any analytical phase function to desired precision.
%Empirically, the use of higher order phase models allow for higher quality
%manipulations.
%
%The technique we will consider in this paper is the Distribution Derivative
%Method for its convenience: the technique does not require computing derivatives
%of the signal to be analysed. It does, however, require a
%$p$ times differentiable analysis window where $p+1$ is the order of the
%phase polynomial to be estimated.
%
%In the few reports on the technique, the window used is the Hann window. This
%window is many-times differentiable but 
%
%Complex polynomial phase models whose estimation is 
\section{The polynomial phase sinusoid}
Recently techniques estimating the parameters of polynomial phase sinusoids,
i.e., signals of the form
\[
    x(t) = \exp(a_0 + \sum_{j=1}^{Q} a_j t^j)
\]
have been developed. Many equivalent techniques have been developed for
estimating the $a_j$. In this paper we compare with the
Distribution Derivative Method (DDM) only, without loss of generality.

\section{The estimation of $a_j$}
We will now show briefly how the DDM can be used to estimate the $a_j$. Based on
the theory of distributions (Laurent Schwartz), the DDM
makes use of ``test
functions'' or atoms $\psi$. These atoms must be once differentiable with respect to
$t$ and be non-zero only on a finite interval $[-\frac{L_{t}}{2},\frac{L_{t}}{2}]$. First, we define
%
\begin{equation}
    \label{eq:ddm:inner:prod:def}
    \left\langle x , \psi \right\rangle = 
    \int_{-\infty}^{\infty}x(t)\overline{\psi}(t)dt
\end{equation}
%
and the operator 
%
\[
\mathcal{T}^{\alpha} : (\mathcal{T}^{\alpha}x)(t) = t^{\alpha}x(t)
\]
%
Consider the weighted signal
%
\[
    f(t) = x(t) \overline{\psi}(t)
\]
%
differentiating with respect to $t$ we obtain
%
\begin{multline}
    \label{eq:ddm:weighted:sig:derivative}
    \frac{df(t)}{dt} = 
    \frac{dx}{dt}(t)\overline{\psi}(t)
    + x(t)\frac{d\overline{\psi}}{dt}(t) \\
    = \left( \sum_{j=1}^{Q} j a_j t^{j-1} \right) x(t)\overline{\psi}(t)
    + x(t)\frac{d\overline{\psi}}{dt}(t)
\end{multline}
%
Because $\psi$ is zero outside of the interval $[-\frac{L_{t}}{2},\frac{L_{t}}{2}]$, integrating
$\frac{df(t)}{dt}$ we obtain
%
\begin{multline*}
    \int_{-\infty}^{\infty}\frac{df(t)}{dt}dt \\
    = \sum_{j=1}^{Q} j a_j \int_{-\frac{L_{t}}{2}}^{\frac{L_{t}}{2}} t^{j-1} x(t) \overline{\psi}(t) dt
    + \left\langle x, \frac{d\overline{\psi}}{dt} \right\rangle = 0
\end{multline*}
%
or, using the operator $\mathcal{T}^{\alpha}$,
%
\[ 
    \sum_{j=1}^{Q} j a_j 
    \left\langle \mathcal{T}^{j-1} x , \overline{\psi} \right\rangle
    = -\left\langle x, \frac{d\overline{\psi}}{dt} \right\rangle
\]
%

@@

Although this technique can be exteded to describe a single
sinusoid of arbitrary complexity simply by increasing $Q$, it is still of
interest to consider a signals featuring a sum of $P$ such components, i.e.,
%
\[
    s(t) = \sum_{i=1}^{P} x_{i}(t)
\]
%
with
%
\[
    x_{i}(t) = \exp(a_{i,0} + \sum_{j=1}^{Q} a_{i,j} t^j)
\]
%
In situations where multiple signal sources are known to exist (e.g., recordings
of multiple speakers or performers) such a model is plausible.

@@

\section{Designing the $\psi_{r}$}

In practice, an approximation of \ref{eq:ddm:inner:prod:def} is evaluated using
the discrete Fourier transform (DFT) on a signal $x$ that is properly sampled
and so can be evaluated at a finite number of points $nT$ with $n \in [0,N-1]$ and
$T$ the sample period in seconds. In this way, the atoms $\psi_{\omega}(t)$ chosen
are the products of the elements of the Fourier basis and an appropriately
chosen window $w$ that is once differentiable and finite, i.e.,
%
\[
    \psi_{\omega}(t) = w(t) \exp(-\sqrt{-1} \omega t)
\]
%
Defining $N = \frac{L_{t}}{T}$ and $\omega = 2
\pi \frac{r}{N}$, the approximate
inner product is then
%
\[
    \left\langle x , \psi_{\omega} \right\rangle \approx 
    \frac{1}{N} \sum_{n=0}^{N-1} x(Tn) w(Tn) \exp(-2 \pi \sqrt{-1} r \frac{n}{N}) 
\]
%
i.e., the definition of the DFT of a windowed signal.
\footnote{%
    Notice however that this is an approximation of the inner product and should
    not be interpreted as yielding the Fourier series coefficients of a properly
    sampled signal $x$ periodic in $L_{t}$. This means that other
    evaluations of the inner product that yield more accurate results are
    possible. For example, the analytic solution is possible if $x$ is assumed
    zero outside of $[-\frac{L_{t}}{2},\frac{L_{t}}{2}]$ (the $\psi$ are in
    general analytic).  In this case the samples of $x$ are convolved with the
    appropriate interpolating sinc functions and the integral of this function's
    product with $\psi$ is evaluated.
}%
The DFT is readily interpreted as a bank of bandpass filters centred at
normalized frequencies $\frac{r}{N}$ and with frequency response described by
the Fourier transform of modulated $w$. Therefore choosing $\psi$ amounts to a
filter design problem under the constraints that the impulse response of the
filter be differentiable in $t$ and finite. To minimize the influence of
all but one component, granted the components's energy concentrations are
sufficiently separated in frequency, we desire impulse responses whose magnitude
response gives maximum
out-of-band rejection or equivalently, windows whose Fourier transform exhibits
the lowest side-lobes.

@@

An differentiable approximation to a symmetrical window can be designed in a
straight-forward way. (Harris) and (McGonnegal) have shown how to design optimal
windows of length $N$ samples using a linear combination of $M$ harmonically
related cosines
\begin{equation}
    w(n) = \sum_{k=0}^{M-1} b_{k} \cos (2 \pi k \frac{n}{N})
\mathcal{R}(\frac{n}{N})
\end{equation}
where $\mathcal{R}$ is the \textit{rectangle function}. This function is discontinous
at $n = \frac{\pm N}{2}$, and therefore not differentiable there, unless
\[
\sum_{k=0}^{M-1} b_{k} \cos ( \pm \pi k ) = 0
\]

We choose the $b_{k}$ so that the window $w$'s squared approximation error to
$v$ is minimized
while having $w(\frac{\pm N}{2}) = 0$, i.e. we find the solution $\{
b^{\ast}_{k} \}$ to the mathematical program
\begin{equation}
    \text{minimize}
    \sum_{n=0}^{N-1} ( v(n) 
        - \sum_{k=0}^{M-1} b_{k} \cos(2 \pi k \frac{n}{N}))^{2}
\end{equation}
\[
    \text{subject to} \\
    \sum_{k=0}^{M-1} b_{k} \cos(\pi k ) = 0
\]
which can be solved using standard numerical linear algebra routines (Goulob and
Van Loan p 585).

\section{Approximating digital prolate spheroidal windows}

Digital prolate spheroidal windows are a parametric approximation to functions
proposed by Slepian et al. maximimzing the energy in a given bandwidth. These can be
tuned to have extremely low side-lobes, at the expense of main-lobe width.
Differentiation of these window functions may be possible but is not as
straightforward as differentiation of the sum-of-cosine windows. Furthermore,
the windows do not generally have end-points equal to 0. In the following we
will demonstrate how to approximate one of these windows.

(Verna, Bilbao et al.) showed how to construct prolate spheroidal windows under
parmeters $N$, the window length in samples, and $W$. We chose $N$ based on the
window length chosen in (Betser) for ease of comparison. $W^{\ast}$ was chosen
by synthesizing windows with $W$ ranging between $0.01$ and $0.99$ at a resolution
of $0.01$. The window with the lowest height of the highest sidelobe was obtained
with $W=0.77$. It is shown in (figure). 
mainlobe width.

The approximation method yields a window virtually identical to the digital
prolate spheroidal window: in our realisation, the approximation error was
everywhere less than $10^{-13}$.

>>>>

\usepackage{hyphenat}
\hyphenation{spher-oidal}

@@

\section{Introduction}
\label{sec:intro}
%Additive synthesis using a sum of sinusoids plus noise is a powerful model for
%representing audio, allowing for the easy implementation of many manipulations
%such as time-stretching and transposition. In its most basic form, peaks in the
%short-time spectrum are identified as representing sinusoids and connected based
%on some criteria. As the phase evolution of the sinusoid is assumed linear over
%the analysis frame, only the phase and frequency of the sinusoids at these
%analysis points are used to fit a plausible phase function. Recently, there has
%been interest in using higher order phase functions as the estimation of their
%parameters has been made possible by a new set of techniques of only moderate
%computational complexity using signal derivatives. Theoretically, the estimation
%of the parameters of a polynomial phase function of arbitrary order make it
%possible to approximate any analytical phase function to desired precision.
%Empirically, the use of higher order phase models allow for higher quality
%manipulations.
%
%The technique we will consider in this paper is the Distribution Derivative
%Method for its convenience: the technique does not require computing derivatives
%of the signal to be analysed. It does, however, require a
%$p$ times differentiable analysis window where $p+1$ is the order of the
%phase polynomial to be estimated.
%
%In the few reports on the technique, the window used is the Hann window. This
%window is many-times differentiable but 
%
%Complex polynomial phase models whose estimation is 
\section{The polynomial phase sinusoid}
Recently techniques for estimating the parameters of complex polynomial phase sinusoids,
i.e., signals of the form
\begin{equation}
    \label{eq:polyphaseexp}
    x(t) = \exp(a_0 + \sum_{j=1}^{Q} a_j t^j)
\end{equation}
with $a_{i} \in \mathbb{C}$, have been developed. Many equivalent techniques have
been developed for estimating the $a_j$. In this paper we compare with the
Distribution Derivative Method (DDM) only, without loss of generality.

\section{The estimation of $a_j$}
We will now show briefly how the DDM can be used to estimate the $a_j$. Based on
the theory of distributions (Laurent Schwartz), the DDM
makes use of ``test
functions'' or atoms $\psi$. These atoms must be once differentiable with respect to
time variable $t$ and be non-zero only on a finite interval $[-\frac{L_{t}}{2},\frac{L_{t}}{2}]$. First, we define
%
\begin{equation}
    \label{eq:ddm:inner:prod:def}
    \left\langle x , \psi \right\rangle = 
    \int_{-\infty}^{\infty}x(t)\overline{\psi}(t)dt
\end{equation}
%
and the operator 
%
\[
\mathcal{T}^{\alpha} : (\mathcal{T}^{\alpha}x)(t) = t^{\alpha}x(t)
\]
%
Consider the weighted signal
%
\[
    f(t) = x(t) \overline{\psi}(t)
\]
%
differentiating with respect to $t$ we obtain
%
\begin{multline}
    \label{eq:ddm:weighted:sig:derivative}
    \frac{df(t)}{dt} = 
    \frac{dx}{dt}(t)\overline{\psi}(t)
    + x(t)\frac{d\overline{\psi}}{dt}(t) \\
    = \left( \sum_{j=1}^{Q} j a_j t^{j-1} \right) x(t)\overline{\psi}(t)
    + x(t)\frac{d\overline{\psi}}{dt}(t)
\end{multline}
%
Because $\psi$ is zero outside of the interval $[-\frac{L_{t}}{2},\frac{L_{t}}{2}]$, integrating
$\frac{df(t)}{dt}$ we obtain
%
\begin{multline*}
    \int_{-\infty}^{\infty}\frac{df(t)}{dt}dt \\
    = \sum_{j=1}^{Q} j a_j \int_{-\frac{L_{t}}{2}}^{\frac{L_{t}}{2}} t^{j-1} x(t) \overline{\psi}(t) dt
    + \left\langle x, \frac{d\overline{\psi}}{dt} \right\rangle = 0
\end{multline*}
%
or, using the operator $\mathcal{T}^{\alpha}$,
%
\[ 
    \sum_{j=1}^{Q} j a_j 
    \left\langle \mathcal{T}^{j-1} x , \overline{\psi} \right\rangle
    = -\left\langle x, \frac{d\overline{\psi}}{dt} \right\rangle
\]
%

@@

Although this technique can be exteded to describe a single
sinusoid of arbitrary complexity simply by increasing $Q$, it is still of
interest to consider a signals featuring a sum of $P$ such components, i.e.,
%
\begin{equation}
    \label{eq:polyphaseexpmix}
    s(t) = \sum_{i=1}^{P} x_{i}(t)
\end{equation}
%
with
%
\[
    x_{i}(t) = \exp(a_{i,0} + \sum_{j=1}^{Q} a_{i,j} t^j)
\]
%
In situations where multiple signal sources are known to exist (e.g., recordings
of multiple speakers or performers) such a model is plausible.

@@

\section{Designing the $\psi_{r}$}
\label{sec:designingatoms}

In practice, an approximation of Eq.~\ref{eq:ddm:inner:prod:def} is evaluated using
the discrete Fourier transform (DFT) on a signal $x$ that is properly sampled
and so can be evaluated at a finite number of points $nT$ with $n \in [0,N-1]$ and
$T$ the sample period in seconds. In this way, the atoms $\psi_{\omega}(t)$ chosen
are the products of the elements of the Fourier basis and an appropriately
chosen window $w$ that is once differentiable and finite, i.e.,
%
\[
    \psi_{\omega}(t) = w(t) \exp(-\sqrt{-1} \omega t)
\]
%
Defining $N = \frac{L_{t}}{T}$ and $\omega = 2
\pi \frac{r}{N}$, the approximate
inner product is then
%
\[
    \left\langle x , \psi_{\omega} \right\rangle \approx 
    \sum_{n=0}^{N-1} x(Tn) w(Tn) \exp(-2 \pi \sqrt{-1} r \frac{n}{N}) 
\]
%
i.e., the definition of the DFT of a windowed signal.%
\footnote{%
    Notice however that this is an approximation of the inner product and should
    not be interpreted as yielding the Fourier series coefficients of a properly
    sampled signal $x$ periodic in $L_{t}$. This means that other
    evaluations of the inner product that yield more accurate results are
    possible. For example, the analytic solution is possible if $x$ is assumed
    zero outside of $[-\frac{L_{t}}{2},\frac{L_{t}}{2}]$ (the $\psi$ are in
    general analytic).  In this case the samples of $x$ are convolved with the
    appropriate interpolating sinc functions and the integral of this function's
    product with $\psi$ is evaluated.
}%
The DFT is readily interpreted as a bank of bandpass filters centred at
normalized frequencies%
\footnote{%
In this article as in much of the literature on signal processing the index $r$
of the Fourier basis element corresponding to the frequency $\omega_{r} = 2 \pi
\frac{r}{N}$ will be referred to as its bin number.
}
$\frac{r}{N}$ and with frequency response described by
the Fourier transform of modulated $w$. Therefore choosing $\psi$ amounts to a
filter design problem under the constraints that the impulse response of the
filter be differentiable in $t$ and finite. To minimize the influence of
all but one component, granted the components's energy concentrations are
sufficiently separated in frequency, we desire impulse responses whose magnitude
response gives maximum
out-of-band rejection or equivalently, windows whose Fourier transform exhibits
the lowest side-lobes.

@@

A differentiable approximation to a symmetrical window can be designed in a
straight-forward way. (Harris) and (McGonnegal) have shown how to design optimal
windows of length $N$ samples using a linear combination of $M$ harmonically
related cosines
\begin{equation}
    w(n) = \sum_{k=0}^{M-1} b_{k} \cos (2 \pi k \frac{n}{N})
\mathcal{R}(\frac{n}{N})
\end{equation}
where $\mathcal{R}$ is the \textit{rectangle function}. This function is
discontinuous
at $n = \frac{\pm N}{2}$, and therefore not differentiable there, unless
\[
\sum_{k=0}^{M-1} b_{k} \cos ( \pm \pi k ) = 0
\]

We choose the $b_{k}$ so that the window $w$'s squared approximation error to
$v$ is minimized
while having $w(\frac{\pm N}{2}) = 0$, i.e. we find the solution $\{
b^{\ast}_{k} \}$ to the mathematical program
\begin{equation}
    \label{eq:searchcontwinprogram}
    \text{minimize}
    \sum_{n=0}^{N-1} ( v(n) 
        - \sum_{k=0}^{M-1} b_{k} \cos(2 \pi k \frac{n}{N}))^{2}
\end{equation}
\[
    \text{subject to} \\
    \sum_{k=0}^{M-1} b_{k} \cos(\pi k ) = 0
\]
which can be solved using constrained least-squares -- a standard numerical
linear algebra routine (Goulob and Van Loan p 585).

\section{Continuous window design example}
\label{sec:designexample}
\begin{figure*}[t]
    \centerline{\includegraphics[width=5in]{{search_dpw_bw_m_fig_mainlobe}.eps}}
\caption{\label{fig:dpwmainlobe} Comparing the mainlobe characteristics of the
continuous 4-term Nuttall window, the digital prolate window with $W=0.008$, and
the continuous approximation to the digital prolate window.}
\end{figure*}

\begin{figure*}[t]
    \centerline{\includegraphics[width=5in]{{search_dpw_bw_m_fig_asymptotic}.eps}}
\caption{\label{fig:dpwasymptotic} Comparing the asymptotic power spectrum
characteristics of the continuous 4-term Nuttall window, the digital prolate
window with $W=0.008$, and the continuous approximation to the digital prolate
window.}
\end{figure*}

As a design example we show how to create a continuous approximation of a digital
prolate spheroidal window.

Digital prolate spheroidal windows are a parametric approximation to functions
proposed by Slepian et al. whose Fourier transform's energy is maximized in a
given bandwidth. These can be tuned to have extremely low side-lobes, at the
expense of main-lobe width.  Differentiation of these window functions may be
possible but is not as straightforward as differentiation of the sum-of-cosine
windows above. Furthermore, the windows do not generally have end-points equal to 0.
In the following we will demonstrate how to approximate a digital prolate
spheroidal window.

(Verna, Bilbao et al.) showed how to construct prolate spheroidal windows under
parameters $N$, the window length in samples, and $W$. We chose $N=512$ based on
the window length chosen in (Betser) for ease of comparison. Its parameter value
was chosen by synthesizing windows with $W$ ranging between $0.005$ and $0.010$
at a resolution of $0.001$. The window with the closest 3db bandwidth to the
4-term Nuttall window was obtained with $W=0.008$. Its magnitude response is
shown in Fig.~\ref{fig:dpwmainlobe}~and~\ref{fig:dpwasymptotic}. We see that
this window's asymptotic roll-off is 6 dB per octave and is therefore not
continuous (Nuttall). We designed an approximate window using
Eq.~\ref{eq:searchcontwinprogram} for $M$ varying between $2$ and $N/8$ to find
the best approximation to the digital prolate window's mainlobe using a small
number of cosines. The $M$ giving the best approximation was $5$. The magnitude
response of the approximation is shown in Fig.~\ref{fig:dpwmainlobe}.  We see
that we obtain a lower highest side-lobe level by slightly sarificing the
narrowness of the mainlobe. Furthermore in Fig.~\ref{fig:dpwasymptotic} we
observe that the rolloff of the window is 18 dB per octave because it is
continuous at all points in its domain.

\section{The performance of improved windows}

\begin{figure*}[ht]
    \centerline{\includegraphics[width=5in]{{ddm_snr_win_comp}.eps}}
\caption{\label{fig:snrwincomp} The estimation variance of random polynomial
phase sinusoids averaged over 1000 trials. \textit{C} is the Cram\'{e}r-Rao
lower bound, \textit{N3} and \textit{N4} are the 3- and 4-cosine-term continuous
Nuttall windows, \textit{H} is the Hann window, and \textit{P5} is the continous
5-cosine-term approximation to a digital prolate window as described in
Sec.~\ref{sec:designexample}.}
\end{figure*}

As in (Betser) we synthesized random sinusoids using Eq.~\ref{eq:polyphaseexp}
with $Q=2$ and compared the average estimation error variance with the
theoretical minimum given by the Cram\'{e}r-Rao bound. The original Hann window,
the windows proposed by Nuttall and the new digital prolate based window were
used to synthesize the atoms as described in Sec.~\ref{sec:designingatoms} and
their their estimation error variance was compared (see
Fig.~\ref{fig:snrwincomp}). After performing the DFT to obtain inner products
with the atoms, three containing the most signal energy were used in the
estimations. We see that windows with lower-sidelobes exhibit the lowest
possible error variance at the sacrifice of some estimator efficiency.

\begin{figure*}[ht]
    \centerline{\includegraphics[width=5in]{{comp_offset_chirp_est_err}.eps}}
\caption{\label{fig:mixwincomp}}
\end{figure*}

To evaluate the performance of the various windows when estimating the
parameters of components in mixture we synthesized signals using
Eq.~\ref{eq:polyphaseexpmix} with $P=2$ and $Q=2$. One of the two components was
scaled to give a desired signal power ratio with the other component. The
parameters of both components were estimated and the total squared estimation
error for all the parameters was recorded. In Fig.~\ref{fig:mixwincomp} we see
that there is generally less estimation error for components having similar
signal power. This is to be expected as the more powerful signal will mask the
other less in these scenarios. The estimation error is large when the atoms
containing the most signal energy for each component are not greatly separated
in frequency. This is likely due to the convolution of the Fourier transform of
the window with the signal -- indeed windows with a larger mainlobe exhibit a
larger `radius' in which the error of the parameter estimation will be high.
However, for atoms sufficiently separated in frequency, windows with lower
sidelobes are better at attenuating the other component and for these the
estimation error is lowest.
